\documentclass[12pt, a4paper]{article} 
\usepackage[usenames,dvipsnames]{xcolor} 
\usepackage[big]{layaureo} 
\usepackage{hyperref} 
\definecolor{linkcolour}{rgb}{0,0.2,0.6}
\hypersetup{colorlinks,breaklinks,urlcolor=linkcolour,linkcolor=linkcolour} 
\usepackage{titlesec}
\titleformat{\section}{\Large\scshape\raggedright}{}{0em}{}[\titlerule]
\titlespacing{\section}{0pt}{3pt}{3pt} 
\pagenumbering{gobble}
\begin{document}

\par{\centering{\Huge James \textsc{Stevenson}}\bigskip\par}

\section{Personal Data}

\large\center 38 Hope Terrace, Edinburgh, EH9 2AR \\
07515927614\\
\href{mailto:james.stevenson09@hotmail.co.uk}{james.stevenson09@hotmail.co.uk}


\section{Profile}

\large{I take pride in my work and always aim to improve my skills.  
Sociable, easy to get on with.
Will pay close attention to set policies and procedures. 
Enthusiastic, interested in people and products. 
Able to communicate ideas clearly. 
Reliable, able to find out the necessary information to excel at any given task.}

\section{Skills and Experience}


\begin{enumerate}
	\item Accurate keyboard skills and proficient in Microsoft Office applications. Familiar with Microsoft Visual Studio.
	\item Capable with object-oriented programming languages such as Java and C\#. 
	\item Can create websites with Bootstrap and HTML. Have used web frameworks like ASP.NET MVC to create web applications. Experience of SQL. Familiar with JavaScript/JQuery.
	\item Familiar with the principles of AGILE software development. Able to use Git for version control.
	\item Experience of using SAP business software. Both SAP ERP and SAP HANA/Predictive Analysis (with R).
	\item Experienced in providing quality customer service in retail and high profile catering events, used to dealing with a wide range of customers.
	\item Working for Splendid Catering and Smart Cuisine required attention to detail in ensuring that customers received their food/drink in a timely fashion. Worked hard in a team to continually improve efficiency.
	\item Involved in training new staff in procedures and customer service at Scotmid and spent time coaching staff at Sothebys in use of Office applications.
	\item I always ensure private information shared by clients, customers or colleagues - both personal and with regards to my workplace - is kept confidential.
\end{enumerate}


\section{Work Experience}
\setlength\tabcolsep{4pt}\centering
\begin{tabular}{r|p{11cm}}

\textsc{Nov 2016 - Jan 2017} & Stockroom Assistant at \textsc{Waterstones}, Edinburgh \\
& \small{Seasonal work to assist with busiest time of year for retail. Used computer systems to locate stock, send invoices and create dispatch notes. Packaged and sent off deliveries within a tight timeframe (5pm Royal Mail pick-up daily). Dealt with large increase of ship-from-shop orders over previous years, often single handedly.} \\
\multicolumn{2}{c}{} \\

\textsc{Aug 2015 - May 2016} & Sales Assistant at \textsc{Scotmid}, Edinburgh \\
& \small{Busy shop environment. Helped customers with their shopping, served on tills. Managed stock, delivery and displays instore. Handling cash and card transactions. Adhering to licensing regulations and training new staff.} \\
\multicolumn{2}{c}{} \\

\textsc{Jan 2015-Aug 2015} & Caterer at \textsc{Smart Cuisine}, Aberdeen \emph{}\\
& \small{Serving customers, taking orders, organising and setting-up events/venues. Required high-standard of professionalism and customer service at corporate and private events.}\\
\multicolumn{2}{c}{} \\

\textsc{Sep 2013 - May 2014} & Caterer at \textsc{Splendid Catering}, Aberdeen \emph{}\\
& \small{Waiting tables, mixing and serving drinks, setting-up and dismantling equipment at venues for high-profile events.}\\
\multicolumn{2}{c}{} \\

\textsc{Aug 2013} & Internship at \textsc{Sotheby's}, Edinburgh \emph{}\\
& \small{Checking lot numbers and cataloging high value stock items accurately. Updating stock records, filing, photocopying. Tidying up old records, and removing old files. Dealing with initial phone enquiries from clients and referring on to the correct department/member of staff. Assisting other members of staff with their IT, coaching staff in Microsoft Office applications, eg. Word and PowerPoint. Assisting CEO with production of his presentation to Edinburgh University.}
\end{tabular}

\section{Education}

\begin{tabular}{rl}	
\textsc{Jan} 2017 - \textsc{Jun} 2018 & MSc in \textsc{Computing} \\
&\textbf{Edinburgh Napier}\\
&\small\emph{In progress}\\
\\

\textsc{Sep} 2015 - \textsc{Sep} 2016 & MSc in \textsc{Philosophy} \\ 
&\textbf{The University of Edinburgh}\\
\\

\textsc{Sep} 2011 - \textsc{Jun} 2015& MA Hons. in \textsc{Philosophy}-\textsc{History of Art}\\ 
&\textbf{The University of Aberdeen}\\ 
&\small\emph{Upper Second Class Honours}\\
\\

\textsc{Aug} 2005 - \textsc{Jun} 2011&\textbf{George Watson's College}, Edinburgh\\
&\small \emph{Adv. Higher:} Economics- B; RMPS- B\\
&\footnotesize \emph{Higher:} English-A; Economics-A; Philosophy-A; Classics-A\\
\end{tabular}


\section{Additional Information}

\begin{tabular}{rl}

\textsc{Driving:} & Holder of full UK driver's licence\\
\textsc{French:} & Conversational\\

\end{tabular}


\end{document}
